Stack Safety as a Security Property

### 2020-07-10 main.tex
CP

But what {\em is} this stack safety of which people speak? Precise formal
definitions are mostly lacking: rather, the notion
is commonly defined in a negative way, by enumerating the bad things that a
given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a
positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as
the conjunction of two properties: {\em local state encapsulation} and {\em
  well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft a more abstract characterization
of stack safety that is independent of capabilities or any other specific
enforcement
technique.  In other words, we seek formal criteria that can be applied
to \emph{any} mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
%
Our key contribution is an abstract
presentation of local state encapsulation using the technical
framework of language-based security, together with a simple
characterization of well-bracketed control flow.

@@

The intuition behind \citeauthor{Skorstengaard+19}'s {\em
  well-bracketed control flow} is the need to protect the call graph
itself: a callee should pass control back to the caller only by
transferring control to the instruction
following the call; it should not be allowed to jump back into the
caller's code at an arbitrary point, either directly (via a jump
instruction) or indirectly (by corrupting stack data and
returning). Well-bracketed control flow establishes safety between
functions: each instruction is uniquely owned by some function, and
all jumps between functions are either calls that enter at a marked
entry point or returns that re-enter just after their associated call
in the standard nested call structure.  \apt{clarify what we are
  contributing here. We are not trying to phrase this as a security
  property.}\leo{Better now? Should this be fixed two paragraphs above
  too?}\bcp{Hmmm this seems like a fundamental issue with the way we are
  framing the story...}

To phrase stack safety, and especially local state encapsulation, as a
security property requires an explicit attacker model: a clear
understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.

>>>>

But what {\em is} this stack safety of which people speak? Precise formal
definitions are mostly lacking: rather, the notion
is commonly defined in a negative way, by enumerating the bad things that a
given enforcement mechanism
prevents, such as corruption of return addresses, buffer overflows, or
use of uninitialized variables.
%\bcp{Does use of unitialized variables lead to stack safety attacks?\apt{sure, via buffer overflow}}
%with the notion usually being
%defined negatively: through examples of stack-based exploits
%\rb{usually though not always; as noted below there are other
%  approaches}\bcp{Yes, we have to be very careful about this}.
%\leo{Examples at a high level}: such as -- does not overwrite return addresses, or ...
Indeed, we are only aware of one quite recent attempt to frame a
positive definition, by
\citet{Skorstengaard+19}, which defines stack safety as
the conjunction of two properties: {\em local state encapsulation} and {\em
  well-bracketed control flow}.
These properties are formalized by defining a capability-based machine
that ``satisfies [stack safety] by construction.''
%
While this work represents an important step in the right direction, we believe
there is still room to craft a more abstract characterization
of stack safety that is independent of capabilities or any other specific
enforcement
technique.  In other words, we seek formal criteria that can be applied
to a mechanism for protecting stack-based low-level programs to judge
whether or not it actually ``guarantees stack safety.''
%
Our key contribution is an abstract
presentation of local state encapsulation using the technical
framework of language-based security, together with a simple
characterization of well-bracketed control flow.

@@

The intuition behind \citeauthor{Skorstengaard+19}'s {\em
  well-bracketed control flow} is the need to protect the call graph
itself: a callee should pass control back to the caller only by
transferring control to the instruction
following the call; it should not be allowed to jump back into the
caller's code at an arbitrary point, either directly (via a jump
instruction) or indirectly (by corrupting stack data and
returning). Well-bracketed control flow establishes safety between
functions: each instruction is uniquely owned by some function, and
all jumps between functions are either calls that enter at a marked
entry point or returns that re-enter just after their associated call
in the standard nested call structure.

To phrase local state encapsulation as a
security property requires an explicit attacker model: a clear
understanding of who is being protected and from what. Most existing
stack protection mechanisms envisage scenarios where a callee is
influenced by external inputs to corrupt the stack in some way. To
avoid getting into the details of exactly how this happens when
formalizing our properties, we use an even stronger model: when we
state the protections that a function can rely on when it makes a
call, we assume that the callee can execute arbitrary instructions
until it returns.
